---
title: "Statistical Inference project Part 2"
output:
  pdf_document: default
  html_document: default
---

## Overview
This report is for second part of the course project of the Coursera course "Statistical Inference" which is a part of specialization "Data Science". In this second part, we perform basic inferential analyses using the ToothGrowth data in the R datasets package.


## 1. Load the ToothGrowth data and perform some basic exploratory data analyses

```{r}
#load the dataset
library(datasets)
data(ToothGrowth)

# check the dataset details
str(ToothGrowth)
```
```{r}
# number of rows of dataset
nrow(ToothGrowth)
```

```{r}
# convert variable dose from numeric to factor
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

# look at the dataset variables after conversion
str(ToothGrowth)
```

## 3. Provide a basic summary of the data

```{r}
# summary statistics for all variables
summary(ToothGrowth)
```

```{r}
#cases per different doses and delivery methods
table(ToothGrowth$dose, ToothGrowth$supp)
```

```{r}
#loading ggplot2
library(ggplot2)
```
```{r}
ggplot(aes(x=dose, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=dose))
```

```{r}
# visualization of tooth growth as function of supplement type
ggplot(aes(x=supp, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=supp))
```

## 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 
```{r}
t.test(len ~ supp, data = ToothGrowth)
```
The confidence interval contains zero which indicates that we can not reject the null hypothesis which states that the different supplement types have no effect on tooth length

```{r}
# first create three sub-groups as per dose level pairs
ToothGrowth.doses_0.5_1.0 <- subset (ToothGrowth, dose %in% c(0.5, 1.0)) 
ToothGrowth.doses_0.5_2.0 <- subset (ToothGrowth, dose %in% c(0.5, 2.0)) 
ToothGrowth.doses_1.0_2.0 <- subset (ToothGrowth, dose %in% c(1.0, 2.0)) 

# Check for group differences due to different dose levels (0.5, 1.0)
# assuming unequal variances between the two groups
t.test(len ~ dose, data = ToothGrowth.doses_0.5_1.0)
```

```{r}
# Check for group differences due to different dose levels (0.5, 2.0)
t.test(len ~ dose, data = ToothGrowth.doses_0.5_2.0)
```
```{r}
# Check for group differences due to different dose levels (1.0, 2.0)
t.test(len ~ dose, data = ToothGrowth.doses_1.0_2.0)
```
For all the 3 subsets, the confidence levels does not contain 0 and mean length increases with increase in dose. We can reject the null hypothesis and put forward that increasing the dose increases in tooth length

## 4. State your conclusions and the assumptions needed for your conclusions.

CONCLUSIONS:
1. supplement type has no impact on tooth growth
2. Dosage impacts tooth length, increase of which increases the tooth length

ASSUMPTIONS:
1. The sample size of guinea pigs is 60, which is used for conclusions
2. For the t-tests, the variances are assumed to be different for the two groups being compared. 




