---
output:
  pdf_document: default
  html_document: default
---
----------------------------------------------------------------------------------
title: "Statistical Inference course project" Athni K
output: html_document
----------------------------------------------------------------------------------

## Overview
This is a project on Statistical Inference. It consists of two parts:
1. A simulation exercise
2. Basic inferential data analysis

This project is to explore inference and do simple inferential data analysis.

##Assignment
In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.


## Simulations
```{r}
#set seed for reproducibility
set.seed(10)

#set lambda with given value
lambda <- 0.2

#set n with given value
n=40

#simulations set to 1000
simulations <-1000

#simulate
sim_exponentials <- replicate(simulations, rexp(n,lambda))

#calculate mean of this
mean_exponentials <- apply(sim_exponentials, 2, mean)
```
##Sample Mean vs Theoretical Mean
```{r}
#distribution mean
samplemean <- mean(mean_exponentials)
samplemean
```
```{r}
#theoretical mean
theoreticalmean <- 1/lambda
theoreticalmean
```
```{r}
#visual representation
hist(mean_exponentials, xlab="mean", main="Exponential function simulation")
abline(v=samplemean, col="red")
abline(v=theoreticalmean, col="blue")
```
The center of distribution of averages of 40 exponentials is very close to the theoretical center of the distribution(5.04 and 5 resp.)


##Sample Variance vs Theoretical Variance
```{r}
#standard deviation of the distribution
samplesd <- sd(mean_exponentials)
samplevariance <- samplesd^2
samplevariance
```

```{r}
#theoretical standard deviation
theoreticalsd <- (1/lambda)/sqrt(n)
theoreticalvariance <- theoreticalsd^2
theoreticalvariance
```
The variances by theory and analytically are very close to each other.


```{r}
xfit <- seq(min(mean_exponentials), max(mean_exponentials), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(n)))
hist(mean_exponentials,breaks=n,prob=T,col="green",xlab = "means",main="Density of means",ylab="density")
lines(xfit, yfit, pch=22, col="black", lty=5)


```

##Distribution
Looking at the distribution, it looks to be a normal distribution


